### 2023-09-21 misc/expo/kelleymeka.tex
Update bio

\proof Expand
$$\hat{f*g}(\chi) = \ex_{x\in G} (f*g)(\chi)\chi(-x)$$
and multiply the right-hand side by $1 = \chi(-y)\chi(y)$ to get
$$\hat{f*g}(\chi) = \ex_{x\in G} \ex_{y\in G} f(y)g(x-y) \chi(-y) \chi(y-x).$$
Then we may interchange the order of summation and substitute $z = x-y$ to arrive at
$$\hat{f*g}(\chi) = \ex_{y\in G} \ex_{z\in G} f(y)g(z) \chi(-y) \chi(-z) = \hat f(\chi)\hat g(\chi),$$
which proves (i). For part (ii), we expand and multiply by the same $1$ to get
$$\widehat{f\circ g}(\chi) = \ex_{x\in G} (f\circ g)(\chi)\chi(-x)
= \ex_{x\in G} \ex_{y\in G} f(y)g(x+y) \chi(y) \chi(-x-y).$$
We again interchange the order of summation; this time substituting $z = x+y$ gives us
$$\eqalign{
\hat{f\circ g}(\chi) &= \ex_{y\in G} \ex_{z\in G} f(y)g(z) \chi(y)\chi(-z) \cr
&= \bar{\ex_{y\in G} f(y)\chi(-y) } \ex_{z\in G} g(z)\chi(-z) \cr
&=  \bar{\hat f(\chi)}\hat g(\chi),\cr
}$$
which is what we wanted.\slug

@@

\edef\knorms{\the\thmcount}
\proclaim Proposition \advthm. Let $G$ be a finite abelian group, let $k\ge 1$ be an integer, and let
$\chi_0$ denote the identity element of the dual group $\hat G$ of $G$. We have the identity
$$\ex_{x\in G} f(x)^k = {\hat f} ^{*p} (\chi_0).$$

@@

\proof We expand
$$\eqalign{
\hat{\mu-1}(\chi_0) &=  \ex_{x\in G} \bigl(\mu(x)-1\bigr) \chi(-x) \cr
&= \ex_{x\in G} \mu(x)\chi(-x) - \ex_{x\in G} \chi(-x).\cr
&= \hat\mu(\chi) - \ex_{x\in G} \chi(-x).\cr
}$$
If $\chi\ne\chi_0$, then the expectation vanishes, and
if $\chi = \chi_0$, then the expection clearly equals $1$, and $\mu(\chi_0) = \norm{\mu}_1 = 1$, so the whole
expression is zero.\slug

@@

\edef\unbalancing{\the\thmcount}
\parenproclaim Lemma~{\advthm} (Unbalancing of spectrally nonnegative functions). Let $\eps\in (0,1)$
and let $\nu : G\to \RR_{\ge 0}$ be a probability measure with $\hat \nu\ge 0$. If $f : G\to \RR$ has
$\hat f\ge 0$ and $\norm{f}_{p(\nu)} \ge \eps$ for some $p\ge 1$, then
$$ \norm{f+1}_{p'(\nu)} \ge 1 + {\eps\over 2}$$
for some $p' = O\bigl(\eps^{-1} \log(\eps^{-1}) p\bigr)$.

\proof Proposition~{\monotonenorm} tells us that $\norm{f}_{p(\mu)}$ is monotonically increasing in $p$, so
without loss of generality we can pick $p$ odd and at least $5$. As usual, we denote the identity in $\hat G$
by $\chi_0$. Using the Fourier inversion formula
and orthogonality of characters as we did in the proof of Proposition~{\knorms}, we observe that
$$\eqalign{
\norm{f}_{p(\nu)}^p &= \ex_{x\in G}
\Bigl(\sum_{\chi\in \hat G} \hat\nu(\chi)\chi(x)\Bigr)\Bigl(\sum_{\chi\in \hat G} \hat f(\chi)\chi(x)\Bigr)^k \cr
&= \sum_{\chi_1\in \hat G} \cdots \sum_{\chi_{k+1}\in \hat G}
\hat f(\chi_1)\cdots\hat f(\chi_k) \hat \nu(\chi_{k+1} )\ex_{x\in G} \chi_1(x)\cdots\chi_{k+1}(x) \cr
&= \sum_{\chi_1\cdots\chi_{k+1} =\chi_0} \hat f(\chi_1) \cdots \hat f(\chi_k) \hat\nu(\chi_{k+1} )\cr
&= \hat\nu * {\hat f}^{*p}(\chi_0). \cr
}$$
Let $P = \{x\in G : f(x)\ge 0\}$ and let $g(x) = \max\{f(x), 0\}$. It is easy to see that
$2g(x) = f(x) + \bigl|f(x)\bigr|$, so
$$\eqalign{
2\langle \one_P, f^p\rangle_\nu &= 2 \ex_{x\in G} \nu(x)\one_P(x)f(x)^p \cr
&= 2\ex_{x\in G} \nu(x) g(x) f(x)^{p-1}  \cr
&= \langle 2g,f^{p-1}\rangle_\nu \cr
&= \ex_{x\in G} \nu(x) f(x)^p + \bigl\langle |f|, f^{p-1}\bigr\rangle_\nu \cr
&= \hat\nu * {\hat f}^{*p}(\chi_0)+ \bigl\langle |f|, |f|^{p-1}\bigr\rangle_\nu, \cr
}$$
where in the last line we used the fact that $f$ is real-valued as well as evenness of $p-1$.
Since the Fourier transforms of $f$ and $\nu$ are both nonnegative, the first term is nonnegative, so
$$\langle \one_P, f^p\rangle_\nu \ge {\bigl\langle |f|, |f|^{p-1}\bigr\rangle_\nu \over 2}
= {\norm{f}_{p(\nu)}^p\over 2} \ge {\eps^p\over 2} .$$
Now let $T = \{x\in P : f(x) \ge 3\eps / 4\}$. Then
$$\eqalign{
\langle \one_{P\setminus T}, f^p \rangle &= \langle \one_P - \one_T, f^p \rangle \cr
&= \langle \one_P , f^p \rangle - \langle \one_T, f^p \rangle \cr
&\ge {\eps\over 2} - \ex_{x\in G} \one_T(x)f(x)^p \cr
&\ge {\eps\over 2} - \ex_{x\in G} 3\eps/4 \cr
&= {\eps\over 4},\cr
}$$
so $\langle \one_T, f^p \rangle_\nu \ge \eps^p/4$ and we have
$$\eqalign{
{\eps\over 4} &\le \langle \one_T,f^p\rangle_\nu \cr
&= \ex_{x\in G} \bigl( \nu(x)^{1/2}\one_T(x)\bigr) \bigl(\nu(x)^{1/2} f(x)^p\bigr) \cr
&= \Bigl( \ex_{x\in G} \nu(x)\one_T(x)^2\Bigr)^{1/2} \Bigl(\ex_{x\in G}\nu(x) f(x)^{2p}\Bigr)^{1/2} \cr
&= \Bigl( \ex_{x\in G} \nu(x)\one_T(x)\Bigr)^{1/2} \Bigl(\ex_{x\in G} \nu(x) |f(x)|^{2p}\Bigr)^{p/(2p)} \cr
&= \nu(T)^{1/2} \norm{f}_{2p(\nu)}^p\cr
}$$
by the Cauchy--Schwarz inequality.

Now if $\norm{f}_{2p(\nu)}^p > 2$, then we could take $p' = 2p$, so assume that this norm is at most $2$.
By the triangle inequality, we have
$$\norm{f}_{2p(\nu)} \le \norm{-1}_{2p(\nu)} + \norm{f+1}_{2p(\nu)} \le 3,$$
hence
$$ \nu(T)^{1/2} 3^p \le {\eps^p\over 4}.$$
Since we assumed that $p\ge 5$, and since $4 < 1024/243$, we have $4^{1/p} < 4/3$ and thus
$$ \nu(T) \ge {\eps^{2p} \over 16 \cdot 3^{2p}} = \biggl( {\eps \over 4^{1/p} \cdot 3} \biggr)^{2p}
> \biggl({\eps\over 4}\biggr)^{2p}.$$
This allows us to bound
$$\eqalign{
\norm{f+1}_{p'(\nu)}
&= \Bigl( \ex_{x\in G} \nu(x) \bigl| f(x)+1\bigr|^{p'}\Bigr)^{1/p'} \cr
&\ge \Bigl( \ex_{x\in G} \nu(x) \one_T(x) \bigl| f(x)+1\bigr|^{p'}\Bigr)^{1/p'} \cr
&\ge \bigl( \nu(T) (1+3\eps/4)^{p'} \bigr)^{1/p'} \cr
&> \biggl( {\eps \over 4} \biggr)^{2p/p'} \biggl(1+{3\over 4}\eps\biggr). \cr
}$$
Now if $p' \ge (8p/\eps)\log(4/\eps) = O\bigl(\eps^{-1} \log(\eps^{-1}) p\bigr)$, then
$\eps/4 \ge (2p/p')\log(4/\eps)$ and thus
$$ -{2p\over p'} \log\biggl({4\over \eps}\biggr) \ge -{\eps\over 4}.$$
Taking $e$ to the power of both sides gives us
$$\biggl({4\over \eps}\biggr) ^{-2p/p'} \ge e^{-\eps/4} \ge 1-{\eps\over 4},$$
and plugging this in above gives the bound
$$\norm{f+1}_{p'(\nu)} > \biggl(1-{\eps\over 4}\biggr)\biggl(1+{3\over 4}\eps\biggr)
= 1+{\eps \over 2} -{3\eps^2\over 16} \ge 1+{\eps\over 2},$$
which is what we needed.\slug

@@

\proof Suppose that (i) fails. Then by Lemma~{\holderlifting} there is some $p = O\bigl(\log(1/\gamma)\bigr)$
such that $\norm{\mu_A\circ\mu_A - 1}_p\ge \eps/ 2$. By the second convolution law (part (ii) of
Proposition~{\convolutions}), $\mu_A\circ\mu_A$ is a nonnegative function, and note also that if
$\nu$ is the uniform measure on $\FF_q^n$, then for any $\chi:\hat{\FF_q^n}\to \CC$,
$$\hat\nu(\chi) = \ex_{x\in G} \nu(x)\chi(-x) = \cases{q^{-n}, & if $\chi$ is the trivial character;
\cr 0, & otherwise.}$$
This implies in particular that $\hat\nu \ge 0$,
so by Lemma~{\unbalancing} applied with $f=\mu_A\circ\mu_A$ and the uniform measure for $\nu$, we find
that $\norm{\mu_A\circ\mu_A}_{p'}\ge 1+\eps/4$ for some
$p' = O\bigl(2\log(2) p\bigr)= O\bigl(\log(1/\gamma)\bigr)$.
Let $C(\eps) = \eps^{-1}\log(\eps^{-1})$. By Corollary~{\deprandcorollary},
there are sets $A_1,A_2\subseteq G$, each of density
$\Omega(\alpha^{2p'+O(C(\eps))})$, such that
$$\langle \mu_{A_1}\circ\mu_{A_2}, \one_S\rangle \ge 1 - \eps/32,$$
where $S = \bigl\{ x \in G:(\mu_A\circ\mu_A)(x)\ge 1+\eps/8\bigr\}$. Feeding $-A_1$, $A_2$, and $S$
into Theorem~{\almostperiodicity} with $\eps/32$ gives us a subspace $V$ of codimension
$$\eqalign{
O\Bigl(\eps^{-2}&\log\bigl(\eps^{-1}\alpha^{-2p'-O(C(\eps))}\bigr)^2
\log\bigl(\alpha^{-2p'-O(C(\eps))}\bigr)^2\Bigr) \cr
&= O\Bigl(\eps^{-2} (2p'+C(\eps))^2\log(1/\alpha)^2
\bigl((2p'+C(\eps))\log(1/\alpha)+\log(\eps^{-1})\bigr)^2
\Bigr)\cr
&=O\Bigl(\eps^{-2}\bigl(\log(1/\gamma)+C(\eps)\bigr)^4\log(1/\alpha)^4\Bigr) \cr
}$$
such that
$$\bigl| \langle \mu_V*\mu_{-A_1}*\mu_{A_2}, \one_S\rangle -
 \langle \mu_{-A_1}*\mu_{A_2}, \one_S\rangle\bigr| \le \eps/32.$$
It is easily checked that $\mu_{-A_1}*\mu_{A_2} = \mu_{A_1}\circ\mu_{A_2}$, so we find that
$$\bigl\langle \mu_V*(\mu_{A_1}\circ\mu_{A_2}), \one_S\bigr\rangle \ge 1-\eps/16.$$
Now we observe that
$$\bignorm{(\mu_{A_1}\circ\mu_{A_2})\circ\mu_{A_2}}_1
= \ex_{z\in G}\mu_{A_1}(z)\ex_{y\in G} \mu_{A_2}(y+z) \ex_{x\in G} \mu_A(x+y) = 1,$$
so that,
$$\eqalign{
\max_{x\in G}\;(\mu_V*\one_A)(x) &= \alpha\max_{x\in G}\;(\mu_V*\mu_A)(x) \cr
&= \alpha \bignorm{(\mu_{A_1}\circ\mu_{A_2})\circ\mu_A}_1 \max_{x\in G}\;(\mu_V*\mu_A)(x)\cr
&\ge \alpha\bigl\langle \mu_V*\mu_A, (\mu_{A_1}\circ\mu_{A_2})\circ\mu_A\bigr\rangle\cr
&= \alpha\bigl\langle \mu_V*\mu_A * (\mu_{A_1}\circ\mu_{A_2}), \mu_A\bigr\rangle\cr
&= \alpha\bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \mu_A\circ\mu_A\bigr\rangle,\cr
}$$
where in the last two lines we have employed the adjoint property
$\langle f,g*h \rangle = \langle h\circ f,g\rangle$, as well as the commutative properties
$f*g = g*f$ and $\langle f,g\rangle = \langle g,f\rangle$,
all of which hold for functions $f,g,h : G\to \RR$. But by the construction of $S$, we have
$\one_S(x)(\mu_A\circ\mu_A)(x) \ge (1+\eps/8)\one_S$, so
$$\eqalign{
\bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \mu_A\circ\mu_A\bigr\rangle
&\ge \bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \one_S(\mu_A\circ\mu_A)\bigr\rangle \cr
&\ge (1+\eps/8)\bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \one_S\bigr\rangle \cr
&\ge (1+\eps/8)(1-\eps/16) \cr
&\ge 1+\eps/32, \cr
}$$
hence we conclude that
$$\max_{x\in G}\;(\one_A*\mu_V)(x) \ge (1+\eps/32)\alpha\noskipslug$$

\parenproclaim Theorem~{\advthm} (Finite field). Let $q$ be a power of an odd prime and let $A$ be
a subset of $G = \FF_q^n$ of cardinality $\alpha q^n$. The number of (possibly trivial)
three-term arithmetic progressions contained in $A$ is at least
$${\alpha^3\over 2}q^{2n-O(\log(1/\alpha)^9)}.$$
Hence if $A\subseteq \FF_q^n$ contains no nontrivial three-term arithmetic progressions, then
$\alpha \le e^{-\Omega(n^{1/9})} = O\bigl(q^{-\Omega(n^{1/9})}\bigr)$.

\proof Let $C = 2\cdot A = \{2a : a\in A\}$, so that $\gamma = |C|/q^n = \alpha$. By Lemma~{\densityincrement}
applied to $A$ and $C$ with parameter $\eps = 1/2$, we find that either
$\langle \mu_A*\mu_A,\mu_C\rangle \ge 1/2$ or there is a subspace $V$ of codimension
$O\bigl(\log(1/alpha)^8\bigr)$ such that $\max_{x\in G}\;(\one_A*\mu_V)(x)\ge (1+\eps/64)\alpha$.

@@

For the last part of the theorem statement,
suppose that $A$ does not contain any nontrivial three-term arithmetic progressions.
Then
$${\alpha^3\over 2}q^{2n-O(\log(1/\alpha)^9)} \le \bigl| \{ (x,y)\in (2\cdot A) \times A : x-y\in A'\}\bigr|
\le |A'| \le \alpha q^n,$$
whence
$$q^n \le {2q^{O(\log(1/\alpha)^9)}\over \alpha^2},$$
and taking $q$th logs of both sides yields
$$n \le \log_q\biggl({2q^{O(\log(1/\alpha)^9)}\over \alpha^2}\biggr) = O\bigl(\log(1/\alpha)^9\bigr).$$
Letting $C$ be the constant implicit in the $O$-notation, we invert this to obtain the bound
$$\alpha\le e^{-n^{1/9}/C^{1/9}} = e^{-\Omega(n^{1/9})},$$
which is what we wanted.\slug

>>>>

\proof Expand
$$\hat{f*g}(\chi) = \ex_{x\in G} (f*g)(\chi)\chi(-x)$$
and multiply the right-hand side by $1 = \chi(-y)\chi(y)$ to get
$$\hat{f*g}(\chi) = \ex_{x\in G} \ex_{y\in G} f(y)g(x-y) \chi(-y) \chi(y-x).$$
Then we may interchange the order of summation and substitute $z = x-y$ to arrive at
$$\hat{f*g}(\chi) = \ex_{y\in G} \ex_{z\in G} f(y)g(z) \chi(-y) \chi(-z) = \hat f(\chi)\hat g(\chi),$$
which proves (i). For part (ii), we expand and multiply by the same $1$ to get
$$\widehat{f\circ g}(\chi) = \ex_{x\in G} (f\circ g)(x)\chi(-x)
= \ex_{x\in G} \ex_{y\in G} f(y)g(x+y) \chi(y) \chi(-x-y).$$
We again interchange the order of summation; this time substituting $z = x+y$ gives us
$$\eqalign{
\hat{f\circ g}(\chi) &= \ex_{y\in G} \ex_{z\in G} f(y)g(z) \chi(y)\chi(-z) \cr
&= \bar{\ex_{y\in G} f(y)\chi(-y) } \ex_{z\in G} g(z)\chi(-z) \cr
&=  \bar{\hat f(\chi)}\hat g(\chi),\cr
}$$
which is what we wanted.\slug

@@

\edef\knorms{\the\thmcount}
\proclaim Proposition \advthm. Let $G$ be a finite abelian group, let $k\ge 1$ be an integer, and let
$\chi_0$ denote the identity element of the dual group $\hat G$ of $G$. We have the identity
$$\ex_{x\in G} f(x)^k = {\hat f} ^{*k} (\chi_0).$$

@@

\proof We expand
$$\eqalign{
\hat{\mu-1}(\chi) &=  \ex_{x\in G} \bigl(\mu(x)-1\bigr) \chi(-x) \cr
&= \ex_{x\in G} \mu(x)\chi(-x) - \ex_{x\in G} \chi(-x).\cr
&= \hat\mu(\chi) - \ex_{x\in G} \chi(-x).\cr
}$$
If $\chi\ne\chi_0$, then the expectation vanishes, and
if $\chi = \chi_0$, then the expection clearly equals $1$, and $\mu(\chi_0) = \norm{\mu}_1 = 1$, so the whole
expression is zero.\slug

@@

\edef\unbalancing{\the\thmcount}
\parenproclaim Lemma~{\advthm} (Unbalancing of spectrally nonnegative functions). Let $\eps\in (0,1)$
and let $\nu : G\to \RR_{\ge 0}$ have $\norm{\mu}_1 = 1$ and $\hat \nu\ge 0$. If $f : G\to \RR$ has
$\hat f\ge 0$ and $\norm{f}_{p(\nu)} \ge \eps$ for some $p\ge 1$, then
$$ \norm{f+1}_{p'(\nu)} \ge 1 + {\eps\over 2}$$
for some $p' = O\bigl(\eps^{-1} \log(\eps^{-1}) p\bigr)$.

\proof Proposition~{\monotonenorm} tells us that $\norm{f}_{p(\mu)}$ is monotonically increasing in $p$, so
without loss of generality we can pick $p$ odd and at least $5$. As usual, we denote the identity in $\hat G$
by $\chi_0$. Using the Fourier inversion formula
and orthogonality of characters as we did in the proof of Proposition~{\knorms}, we observe that
$$\eqalign{
\norm{f}_{p(\nu)}^p &= \ex_{x\in G}
\Bigl(\sum_{\chi\in \hat G} \hat\nu(\chi)\chi(x)\Bigr)\Bigl(\sum_{\chi\in \hat G} \hat f(\chi)\chi(x)\Bigr)^p \cr
&= \sum_{\chi_1\in \hat G} \cdots \sum_{\chi_{p+1}\in \hat G}
\hat f(\chi_1)\cdots\hat f(\chi_p) \hat \nu(\chi_{p+1} )\ex_{x\in G} \chi_1(x)\cdots\chi_{p+1}(x) \cr
&= \sum_{\chi_1\cdots\chi_{p+1} =\chi_0} \hat f(\chi_1) \cdots \hat f(\chi_p) \hat\nu(\chi_{p+1} )\cr
&= \hat\nu * {\hat f}^{*p}(\chi_0). \cr
}$$
Let $P = \{x\in G : f(x)\ge 0\}$ and let $g(x) = \max\{f(x), 0\}$. It is easy to see that
$2g(x) = f(x) + \bigl|f(x)\bigr|$, so
$$\eqalign{
2\langle \one_P, f^p\rangle_\nu &= 2 \ex_{x\in G} \nu(x)\one_P(x)f(x)^p \cr
&= 2\ex_{x\in G} \nu(x) g(x) f(x)^{p-1}  \cr
&= \langle 2g,f^{p-1}\rangle_\nu \cr
&= \ex_{x\in G} \nu(x) f(x)^p + \bigl\langle |f|, f^{p-1}\bigr\rangle_\nu \cr
&= \hat\nu * {\hat f}^{*p}(\chi_0)+ \bigl\langle |f|, |f|^{p-1}\bigr\rangle_\nu, \cr
}$$
where in the last line we used the fact that $f$ is real-valued as well as evenness of $p-1$.
Since the Fourier transforms of $f$ and $\nu$ are both nonnegative, the first term is nonnegative, so
$$\langle \one_P, f^p\rangle_\nu \ge {\bigl\langle |f|, |f|^{p-1}\bigr\rangle_\nu \over 2}
= {\norm{f}_{p(\nu)}^p\over 2} \ge {\eps^p\over 2} .$$
Now let $T = \{x\in P : f(x) \ge 3\eps / 4\}$. Then
$$\eqalign{
\langle \one_T, f^p \rangle_\nu &= \langle \one_P - \one_{P\setminus T}, f^p \rangle_\nu \cr
&\ge \langle\one_P, f^p\rangle_\nu - \langle \one_{P\setminus T}, f^p\rangle_\nu \cr
&\ge {\eps^p\over 2}- \ex_{x\in G} \one_{P\setminus T} f(x)^p\nu(x)\cr
&> {\eps^p\over 2}- \ex_{x\in G} (3\eps/4)^p \nu(x)\cr
&\ge {\eps^p\over 4}, \cr
}$$
where in the last line we used the fact that $(3/4)^p \le (3/4)^5 < 243/1024 < 4$.
From this we deduce
$$\eqalign{
{\eps^p\over 4} &\le \langle \one_T,f^p\rangle_\nu \cr
&= \ex_{x\in G} \bigl( \nu(x)^{1/2}\one_T(x)\bigr) \bigl(\nu(x)^{1/2} f(x)^p\bigr) \cr
&\le \Bigl( \ex_{x\in G} \nu(x)\one_T(x)^2\Bigr)^{1/2} \Bigl(\ex_{x\in G}\nu(x) f(x)^{2p}\Bigr)^{1/2} \cr
&= \Bigl( \ex_{x\in G} \nu(x)\one_T(x)\Bigr)^{1/2} \Bigl(\ex_{x\in G} \nu(x) |f(x)|^{2p}\Bigr)^{p/(2p)} \cr
&= \nu(T)^{1/2} \norm{f}_{2p(\nu)}^p\cr
}$$
by the Cauchy--Schwarz inequality.

Now if $\norm{f+1}_{2p(\nu)} > 2$, then we could take $p' = 2p$, so assume that this norm is at most $2$.
By the triangle inequality, we have
$$\norm{f}_{2p(\nu)} \le \norm{-1}_{2p(\nu)} + \norm{f+1}_{2p(\nu)} \le 3,$$
hence
$$ \nu(T)^{1/2} 3^p \ge {\eps^p\over 4}.$$
Once again using the fact that $4 < 1024/243$, we have $4^{1/p} < 4/3$ and thus
$$ \nu(T) \ge {\eps^{2p} \over 16 \cdot 3^{2p}} = \biggl( {\eps \over 4^{1/p} \cdot 3} \biggr)^{2p}
> \biggl({\eps\over 4}\biggr)^{2p}.$$
This allows us to bound
$$\eqalign{
\norm{f+1}_{p'(\nu)}
&= \Bigl( \ex_{x\in G} \nu(x) \bigl| f(x)+1\bigr|^{p'}\Bigr)^{1/p'} \cr
&\ge \Bigl( \ex_{x\in G} \nu(x) \one_T(x) \bigl| f(x)+1\bigr|^{p'}\Bigr)^{1/p'} \cr
&\ge \bigl( \nu(T) (1+3\eps/4)^{p'} \bigr)^{1/p'} \cr
&> \biggl( {\eps \over 4} \biggr)^{2p/p'} \biggl(1+{3\over 4}\eps\biggr). \cr
}$$
Now if $p' \ge (8p/\eps)\log(4/\eps) = O\bigl(\eps^{-1} \log(\eps^{-1}) p\bigr)$, then
$\eps/4 \ge (2p/p')\log(4/\eps)$ and thus
$$ -{2p\over p'} \log\biggl({4\over \eps}\biggr) \ge -{\eps\over 4}.$$
Taking $e$ to the power of both sides gives us
$$\biggl({4\over \eps}\biggr) ^{-2p/p'} \ge e^{-\eps/4} \ge 1-{\eps\over 4},$$
and plugging this in above gives the bound
$$\norm{f+1}_{p'(\nu)} > \biggl(1-{\eps\over 4}\biggr)\biggl(1+{3\over 4}\eps\biggr)
= 1+{\eps \over 2} -{3\eps^2\over 16} \ge 1+{\eps\over 2},$$
which is what we needed.\slug

@@

\proof Suppose that (i) fails. Then by Lemma~{\holderlifting} there is some $p = O\bigl(\log(1/\gamma)\bigr)$
such that $\norm{\mu_A\circ\mu_A - 1}_p\ge \eps/ 2$. By the second convolution law (part (ii) of
Proposition~{\convolutions}), $\mu_A\circ\mu_A$ is a nonnegative function, and note also that if
$\nu$ is the uniform measure on $\FF_q^n$, then for any $\chi:\hat{\FF_q^n}\to \CC$,
$$\hat\nu(\chi) = \ex_{x\in G} \nu(x)\chi(-x) = \cases{q^{-n}, & if $\chi$ is the trivial character;
\cr 0, & otherwise.}$$
This implies in particular that $\hat\nu \ge 0$,
so by Lemma~{\unbalancing} applied with $f=\mu_A\circ\mu_A$ and the uniform measure for $\nu$, we find
that $\norm{\mu_A\circ\mu_A}_{p'}\ge 1+\eps/4$ for some
$p' = O\bigl((2/\eps)\log(2/\eps) p\bigr)= O\bigl(\eps^{-1}\log(\eps^{-1})\log(1/\gamma)\bigr)$.
Let $C(\eps) = \eps^{-1}\log(\eps^{-1})$. By Corollary~{\deprandcorollary},
there are sets $A_1,A_2\subseteq G$, each of density
$\Omega(\alpha^{2p'+O(C(\eps))})$, such that
$$\langle \mu_{A_1}\circ\mu_{A_2}, \one_S\rangle \ge 1 - \eps/32,$$
where $S = \bigl\{ x \in G:(\mu_A\circ\mu_A)(x)\ge 1+\eps/8\bigr\}$. Feeding $-A_1$, $A_2$, and $S$
into Theorem~{\almostperiodicity} with $\eps/32$ gives us a subspace $V$ of codimension
$$\eqalign{
O\Bigl(\eps^{-2}&\log\bigl(\eps^{-1}\alpha^{-2p'-O(C(\eps))}\bigr)^2
\log\bigl(\alpha^{-2p'-O(C(\eps))}\bigr)^2\Bigr) \cr
&= O\Bigl(\eps^{-2} (2p'+C(\eps))^2\log(1/\alpha)^2
\bigl((2p'+C(\eps))\log(1/\alpha)+\log(\eps^{-1})\bigr)^2
\Bigr)\cr
&=O\Bigl(\eps^{-2}\bigl(\log(1/\gamma)+C(\eps)\bigr)^4\log(1/\alpha)^4\Bigr) \cr
}$$
such that
$$\bigl| \langle \mu_V*\mu_{-A_1}*\mu_{A_2}, \one_S\rangle -
 \langle \mu_{-A_1}*\mu_{A_2}, \one_S\rangle\bigr| \le \eps/32.$$
It is easily checked that $\mu_{-A_1}*\mu_{A_2} = \mu_{A_1}\circ\mu_{A_2}$, so we find that
$$\bigl\langle \mu_V*(\mu_{A_1}\circ\mu_{A_2}), \one_S\bigr\rangle \ge 1-\eps/16.$$
Now we observe that
$$\bignorm{(\mu_{A_1}\circ\mu_{A_2})\circ\mu_{A_2}}_1
= \ex_{z\in G}\mu_{A_1}(z)\ex_{y\in G} \mu_{A_2}(y+z) \ex_{x\in G} \mu_A(x+y) = 1,$$
so that,
$$\eqalign{
\max_{x\in G}\;(\mu_V*\one_A)(x) &= \alpha\max_{x\in G}\;(\mu_V*\mu_A)(x) \cr
&= \alpha \bignorm{(\mu_{A_1}\circ\mu_{A_2})\circ\mu_A}_1 \max_{x\in G}\;(\mu_V*\mu_A)(x)\cr
&\ge \alpha\bigl\langle \mu_V*\mu_A, (\mu_{A_1}\circ\mu_{A_2})\circ\mu_A\bigr\rangle\cr
&= \alpha\bigl\langle \mu_V*\mu_A * (\mu_{A_1}\circ\mu_{A_2}), \mu_A\bigr\rangle\cr
&= \alpha\bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \mu_A\circ\mu_A\bigr\rangle,\cr
}$$
where in the last two lines we have employed the adjoint property
$\langle f,g*h \rangle = \langle h\circ f,g\rangle$, as well as the commutative properties
$f*g = g*f$ and $\langle f,g\rangle = \langle g,f\rangle$,
all of which hold for functions $f,g,h : G\to \RR$. But by the construction of $S$, we have
$\one_S(x)(\mu_A\circ\mu_A)(x) \ge (1+\eps/8)\one_S$, so
$$\eqalign{
\bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \mu_A\circ\mu_A\bigr\rangle
&\ge \bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \one_S(\mu_A\circ\mu_A)\bigr\rangle \cr
&\ge (1+\eps/8)\bigl\langle \mu_V* (\mu_{A_1}\circ\mu_{A_2}), \one_S\bigr\rangle \cr
&\ge (1+\eps/8)(1-\eps/16) \cr
&\ge 1+\eps/32, \cr
}$$
hence we conclude that
$$\max_{x\in G}\;(\one_A*\mu_V)(x) \ge (1+\eps/32)\alpha\noskipslug$$

\parenproclaim Theorem~{\advthm} (Finite field). Let $q$ be a power of an odd prime and let $A$ be
a subset of $G = \FF_q^n$ of cardinality $\alpha q^n$. The number of (possibly trivial)
three-term arithmetic progressions contained in $A$ is at least
$${\alpha^3\over 2}q^{2n-O(\log(1/\alpha)^9)}.$$
Hence if $A\subseteq \FF_q^n$ contains no nontrivial three-term arithmetic progressions, then
$\alpha \le q^{-\Omega(n^{1/9})}$.

\proof Let $C = 2\cdot A = \{2a : a\in A\}$, so that $\gamma = |C|/q^n = \alpha$. By Lemma~{\densityincrement}
applied to $A$ and $C$ with parameter $\eps = 1/2$, we find that either
$\langle \mu_A*\mu_A,\mu_C\rangle \ge 1/2$ or there is a subspace $V$ of codimension
$O\bigl(\log(1/\alpha)^8\bigr)$ such that $\max_{x\in G}\;(\one_A*\mu_V)(x)\ge (1+\eps/64)\alpha$.

@@

For the last part of the theorem statement,
suppose that $A$ does not contain any nontrivial three-term arithmetic progressions.
Then
$${\alpha^3\over 2}q^{2n-O(\log(1/\alpha)^9)} \le \bigl| \{ (x,y)\in (2\cdot A) \times A : x-y\in A'\}\bigr|
\le |A'| \le \alpha q^n,$$
whence
$$q^n \le {2q^{O(\log(1/\alpha)^9)}\over \alpha^2},$$
and taking $q$th logs of both sides yields
$$n \le \log_q\biggl({2q^{O(\log(1/\alpha)^9)}\over \alpha^2}\biggr) = O\bigl(\log(1/\alpha)^9\bigr)
= O\bigl(\log_q(1/\alpha)^9\bigr).$$
Letting $C$ be the constant implicit in the last big-$O$ bound, we invert this to obtain
$$\alpha\le q^{-n^{1/9}/C^{1/9}} = q^{-\Omega(n^{1/9})},$$
which is what we wanted.\slug
